Storing messages with multipartite neural cliques

### 2014-07-11 Articles/article.tex
...

	 \begin{abstract}
	 	We extend recently introduced associative memories based on clustered cliques to multipartite cliques. We propose a variant of the classic retrieving rule. We study its performance relatively to the former one for retrieving partially erased or corrupted messages. We provide both analytical and simulation results showing improvements in both networks capacities and resilience. We compute asymptotic capacities of these networks.
	 \end{abstract}

@@

	Storing a message $x$ in the network amounts to consider the network in state $y$, that is for all $1 \le j \le l$ and $1 \le i \le c$, the unit $j$ in the cluster $i$ is activated if and only if $y_{i,j} = 1$ ; then adding all connections between activated units excepting for units sharing the same cluster. Thus we obtain a multipartite clique between activated units representing a message. The parameter $a$ describing the number of activated units per cluster will be called number of activities.

@@

		Retrieving messages occurs by message passing.

		\subsubsection{The "winner takes all" rule}

		take only maximum score

		\subsubsection{The "$w$-winners take all" rule}

@@

		Although it is a less biologically plausible rule, it has the same algorithmic complexity, since selecting the $w$-th element in an array of $n$ elements is $O(n)$ in complexity.

@@

	Suppose we are trying to recover a message where $c_e$ clusters have been erased.

@@

	The network is a binary symmetric matrix with null diagonal and can therefore be stored in memory as $\frac{c(c-1) l^2}{2}$ bits. Since messages are not stored in an ordered fashion, the information stored is the set of messages, thereby it amounts to $\log_2(\frac{({l \choose a}^c)^m}{m!}) \mathop{=}_{m \rightarrow +\infty} m(c \log_2{l \choose a } - \log_2(m) + 1 + o(1)) $. The information effectively stored in the network is $P_{retrieve} \times m  \frac{2 \left(c \log_2{l \choose a } - \log_2(m) + 1 \right)}{c(c-1)l^2}$.

@@

	$\eta_m = P_{retrieve}  \frac{2 m\left(c \log_2{l \choose a } - \log_2(m) + 1 \right)}{c(c-1)l^2} \times (1-(\frac{1}{{l \choose a}})^{c - c_e})^{-(m-1)}$
	%\frac{c}{(c-c_e) \log_2{l \choose a }}

@@

	\begin{align*}
	&P_{retrieve} = \left [ \sum_{n = 1}^{a(c-1) + \gamma} \left [ \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = n)^k P(n_{agc} > n)^{a-k} \right ] P(n_{bc} < n)^{l-a} \right ]^{c - c_e}\\ 
	 &\times \left [ \sum_{n = 1}^{a(c-1) +\gamma} \left[ \sum_{k = 1}^a { a \choose k } P(n_{ge} = n)^k P(n_{ge} > n)^{a-k} \right ] P(n_{be} < n)^{l-2a} P(n_{abe} < n)^a \right ]^{c_e} \numberthis \label{from_err_th}
	\end{align*}

@@

	\newpage
	$\,$
	\newpage

@@

		$\eta_m = P_{retrieve}  \frac{2 m\left(c \log_2{l \choose a } - \log_2(m) + 1 \right)}{c(c-1)l^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{l \choose a}} \right )^k \left (\frac{1}{{l \choose a}} \right )^{c-k} \right ] ^{-(m-1)}$

@@

@@

		\includegraphics[scale=0.50]{Courbes/5portant_psi_c8l256e4}
		\caption{Comparison of efficiencies between different number of activities per cluster :  8 clusters, 256 units per cluster, 1000 sampled messages, pools of 5 networks \label{comppsi}}
	\end{figure}

@@

	$\eta_{m, \psi} = P_{retrieve} m  \frac{2 \left(c \log_2{l \choose a } - \log_2(m) + 1 \right)}{c(c-1)l^2} \times (1-{l \choose a}^{c - c_e})^{-(m-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $

@@

		The entropy of the messages set can be approximated as $H(\mathcal{M}) \approx m c \log_2 {l \choose a}$. From $d = 1 - \left (1- \left(\frac{a}{l}\right)^2 \right)^m$, we can deduce $m \sim \frac{\log_2(1-d)}{\log_2(1-(\frac{a}{l})^2)} \sim - \frac{l^2 \log_2(1-d) \log 2}{a^2}$.
		We set $\log_2 {l \choose a} = k c$ (otherwise density approaches $0$ or $1$, and efficiency $0$). It follows that $\eta \sim c \log_2 {l \choose a} \log2(1-d) \sim \frac{k c^2 \log_2(1-d)\log 2}{a^2 {c \choose 2}}$.

		$\eta \sim - 2 k  \log_2 ((1-d)a^{-2}\log 2$.

		We can bound the probability $P_{\exists e}$ for a message being "present" in the network, that is to mean adding it will not add any edge to the network. By union-bound $P_{\exists e} \le {l \choose a}^c d^{a^2 {c \choose 2}} = 2^{c \log_ 2 {l \choose a} + a^2 {c \choose 2 }\log_2(d)} \approx 2^{c^2(k + \frac{a^2}{2} \log_2 (d))}$.

>>>>

	 \begin{abstract}
	 	We extend recently introduced associative memories based on clustered cliques to multipartite cliques. We propose a variant of the classic retrieving rule. We study its performance relatively to the former one for retrieving partially erased or corrupted messages. We provide both analytical and simulation results showing improvements in both networks efficiencies and resilience to damages. We compute asymptotic capacities of these networks.
	 \end{abstract}

@@

	Storing a message $x$ in the network amounts to consider the network in state $y$, that is for all $i$ and $j$ such as $1 \le j \le l$ and $1 \le i \le c$, the unit $j$ in the cluster $i$ is activated if and only if $y_i[j] = 1$ ; then adding all connections between activated units excepting for units sharing the same cluster. Thus we obtain a multipartite clique between activated units representing a message. The parameter $a$ describing the number of activated units per cluster will be called number of activities.

@@

		Retrieving messages occurs by message passing. We adapt the rule first developed in \cite{GriBer20117} to multipartite cliques.

%		\subsubsection{The "winner takes all" rule}
%		
%		take only maximum score
%	
%		\subsubsection{The "$w$-winners take all" rule}

@@

		Although it is a less biologically plausible rule, it has the same algorithmic complexity, since selecting the $w$-th greater element in an array of $n$ elements is $O(n)$ in complexity.

@@

	Suppose we are trying to recover a message where $c_e$ clusters have been erased (i.e. all units in those clusters are set to $0$).

@@

	The network is a binary symmetric matrix with null diagonal and can therefore be stored in memory as $\frac{c(c-1) l^2}{2}$ bits. Since messages are not stored in an ordered fashion, the information stored is the set of messages, thereby it amounts to $\log_2(\frac{({l \choose a}^c)^m}{m!}) \mathop{=}_{m \rightarrow +\infty} m(c \log_2{l \choose a } - \log_2(m) + \frac{1}{\log 2} + o(1)) $ (thanks to  Stirling's formula). The information effectively stored in the network is $P_{retrieve} \times  \frac{2m \left(c \log_2{l \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2}$.

@@

	$\eta_m = P_{retrieve}  \frac{2 m\left(c \log_2{l \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-(\frac{1}{{l \choose a}})^{c - c_e})^{-(m-1)}$
	%\frac{c}{(c-c_e) \log_2{l \choose a }}

@@

	The probability for all good neurons of a correct cluster (resp. of an erroneous cluster) to achieve a score equal or greater than $x$, with at least one having a score of $x$ is $P_c (x) = \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = x)^k P(n_{agc} > x)^{a-k}$ (resp. $P_e(x) = \sum_{k = 1}^a { a \choose k } P(n_{ge} = n)^k P(n_{ge} > n)^{a-k}$)

%	\begin{align*}
%	&P_{retrieve} = \left [ \sum_{n = 1}^{a(c-1) + \gamma} \left [ \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = n)^k P(n_{agc} > n)^{a-k} \right ] P(n_{bc} < n)^{l-a} \right ]^{c - c_e}\\ 
%	 &\times \left [ \sum_{n = 1}^{a(c-1) +\gamma} \left[ \sum_{k = 1}^a { a \choose k } P(n_{ge} = n)^k P(n_{ge} > n)^{a-k} \right ] P(n_{be} < n)^{l-2a} P(n_{abe} < n)^a \right ]^{c_e} \numberthis \label{from_err_th}
%	\end{align*}
		\begin{align*}
	&P_{retrieve} = \left [ \sum_{x = 1}^{a(c-1) + \gamma} P_g(x) P(n_{bc} < x)^{l-a} \right ]^{c - c_e}\\ 
	 &\times \left [ \sum_{x = 1}^{a(c-1) +\gamma} P_e(x) P(n_{be} < x)^{l-2a} P(n_{abe} < x)^a \right ]^{c_e} \numberthis \label{from_err_th}
	\end{align*}

@@

@@

		$\eta_m = P_{retrieve}  \frac{2 m\left(c \log_2{l \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{l \choose a}} \right )^k \left (\frac{1}{{l \choose a}} \right )^{c-k} \right ] ^{-(m-1)}$

@@

	\newpage

@@

		\includegraphics[scale=0.50]{Courbes/5portant_psi_c8l256e4}
		\caption{Comparison of evolutions of efficiency with increasing damages between different number of activities per cluster :  8 clusters, 256 units per cluster%, 1000 sampled messages, pools of 5 networks 
		\label{comppsi}}
	\end{figure}

@@

	$\eta_{m, \psi} = P_{retrieve} m  \frac{2 \left(c \log_2{l \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-{l \choose a}^{c - c_e})^{-(m-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $

@@

		The entropy of the messages set can be approximated as $H(\mathcal{M}) \approx m c \log_2 {l \choose a}$. From $d = 1 - \left (1- \left(\frac{a}{l}\right)^2 \right)^m$, we can deduce $m \sim \frac{\log_2(1-d)}{\log_2(1-(\frac{a}{l})^2)} \sim - \frac{l^2 \log_2(1-d) \log 2}{a^2}$.
		We set $\log_2 {l \choose a} = k c$ (otherwise density approaches $0$ or $1$, and efficiency $0$). It follows that $\eta \sim \frac{m c \log_2 {l \choose a}}{l^2 {c \choose 2}} \sim \frac{k c^2 \log_2(1-d)\log 2}{a^2 {c \choose 2}}$.

		$\eta \sim - 2 k  a^{-2}\log 2 \times \log_2 (1-d)$.

		We can bound the probability $P_{\exists e}$ for a message being "wrongly present" in the network, that is to mean adding it will not add any edge to the network albeit it is not in $\mathcal{M}$. By union-bound $P_{\exists e} \le {l \choose a}^c d^{a^2 {c \choose 2}} = 2^{c \log_ 2 {l \choose a} + a^2 {c \choose 2 }\log_2(d)} \approx 2^{c^2(k + \frac{a^2}{2} \log_2 (d))}$.
